\section{Related Work}
\label{sec:related}

\paragraph{Domain-specific LLM adaptation.}
Fine-tuning LLMs for specialized domains has shown promising results in medicine~\cite{singhal2023medpalm,wu2024pmcllama}, law~\cite{cui2023chatlaw}, and code~\cite{roziere2023codellama}. A common finding is that domain adaptation requires high-quality, domain-specific training data. For niche technical fields like reliability engineering, no such datasets or benchmarks exist, motivating our data generation pipeline.

\paragraph{Self-Instruct and synthetic data.}
Self-Instruct~\cite{wang2023selfinstruct} bootstraps instruction-following data from a small seed set, with extensions including Alpaca~\cite{taori2023alpaca}, WizardLM~\cite{xu2023wizardlm}, and Self-Refine~\cite{ranaldi2024selfrefine}. For verifiable tasks, answer-consistency filtering~\cite{wang2023selfinstruct} has been proposed, but we show in Section~\ref{sec:data:selfinst} that same-model consistency is trivially satisfied in technical domains. SearchInstruct~\cite{li2025searchinstruct} identifies that self-instruct ``depends solely on internal knowledge,'' motivating external grounding. We address correctness through \emph{cross-model} verification (Section~\ref{sec:data:verification}).

\paragraph{Data augmentation for reasoning.}
MetaMath~\cite{yu2024metamath} showed that rephrasing questions from multiple perspectives improves reasoning more than increasing difficulty, with PersonaMath~\cite{wang2024personamath} extending this via persona-driven variants. Shumailov~\etal~\cite{shumailov2024curse} formalized ``model collapse'' risks when training on generated data. Our paraphrase strategy (Section~\ref{sec:data:augment}) is inspired by MetaMath with cross-model verification to prevent quality degradation.

\paragraph{Reinforcement learning for reasoning.}
RLHF~\cite{ouyang2022training} and DPO~\cite{rafailov2023dpo} align models with preferences but rely on noisy preference data. For verifiable tasks, GRPO~\cite{shao2024deepseekmath} generates multiple responses and computes group-relative advantages without a critic model. DeepSeek-R1~\cite{deepseek2025r1} scaled this to elicit sophisticated reasoning. DAPO~\cite{yu2025dapo} introduced dynamic sampling---discarding prompts with zero reward variance---which is critical for small datasets. We apply GRPO with DAPO-style sampling to reliability engineering (Section~\ref{sec:grpo}), showing that SFT warm-start is essential for RL effectiveness.

\paragraph{Parameter-efficient fine-tuning.}
We use 4-bit QLoRA~\cite{dettmers2023qlora,hu2022lora} with NEFTune~\cite{jain2023neftune} regularization, applied to all attention and MLP layers.
